A Comparative Quality Metric for Fuzzing with Logic State Coverage

### 2024-03-03 paper/main.tex
paper: pass

  In this work, we propose \emph{logic state coverage} as a proxy metric to
  count observed interesting behaviors. A logic state distinguishes interesting
  behaviors in a finer granularity than the less interesting ones, making the
  amount of logic state coverage (i.e., observed logic states) directly
  proportional to the number of observed interesting behaviors.
  % 
  We implemented logic state coverage using a bloom filter and reevaluated
  \textsc{FuzzBench} and indecisive-to-subjective fuzzer comparisons in
  hypervisor and grammar fuzzing. 
  % TODO: (The result summary will come here)

@@

As it is challenging to directly count interesting program behaviors when
fuzzing produces hundreds of them every second, we propose \emph{logic state
coverage} as a feasible proxy metric of interesting behaviors. A logic state is
a set of all satisfied branches during one execution, where logic state coverage
is defined as the \emph{individual count} of all observed distinct logic states.
One logic state represents \emph{fewer} behaviors if they are \emph{more
interesting} (i.e., likely relating to unknown abnormal behaviors), meaning that
observing more interesting behaviors is strongly reflected on larger logic state
coverage.

@@

\pp{New vs. unknown behaviors}
%
We first make a clear distinction between \emph{new} and \emph{unknown} 
behaviors. \emph{New} behaviors are, as the name suggests, simply the behaviors 
that have not been observed until now. On the other hand, \emph{unknown}
behaviors are a subset of the \emph{new} behaviors \emph{not equivalent} to any
observed behaviors. In particular, an abnormal behavior is \emph{unknown} if it
is not equivalent to any observed abnormal behaviors (e.g., a unique root
cause). Formally, if $\mathrm{U}$ and $\mathrm{N}$ are the sets of
\underline{\textbf{u}}nknown and \underline{\textbf{n}}ew behaviors,
respectively, $\mathrm{U} \subset \mathrm{N}$.

\pp{Formal justification}
%
The approach of untargeted fuzzing can be described formally as follows. Upon
executing a mutated input, let $P(\mathrm{N})$ be the probability of observing a
\underline{\textbf{n}}ew behavior as a result of execution. Then, the
probability of discovering \underline{\textbf{u}}nknown
\underline{\textbf{a}}bnormal behavior $P(\mathrm{UA}) := P(\mathrm{U} \cap
\mathrm{A})$ is
%
\begin{align}
  P(\mathrm{UA}) 
      &= P(\mathrm{N} \cap \mathrm{UA}) \tag{by $\mathrm{U} \subset \mathrm{N}$} \\
      &= P(\mathrm{N}) \times P(\mathrm{UA} \mid \mathrm{N}) \tag{by chain rule}. 
\end{align}

@@

\pp{Interestingness in practice}
%
As the interestingness $P(\mathrm{UA} \mid \mathrm{N})$ is the expectation of an
\emph{unobserved} behavior, there is no direct way to measure interestingness. 
%
However, the practice of untargeted fuzzing suggests that
the resulting program behavior is unlikely to be \emph{unknown abnormal} if the
mutation base is highly repetitive, as it commonly incorporates corpus
minimization \cite{todo} and branch hit count buckets \cite{todo}.
%
The rationales are: (i) abnormal behaviors might have been already observed with
fewer branch repetitions (i.e., unlikely \emph{abnormal}) and (ii) if so, it may
correspond to pre-observed less-repetition counterparts (i.e., unlikely
\emph{unknown}).

@@

\pp{Logic state coverage}
%
\emph{Logic state coverage} is \emph{the count of observed distinctive logic
states during fuzzing}, where a logic state is called \emph{covered} when
fuzzing observed a program behavior from the logic state (i.e., triggering all
and only the branches in the logic state). 
%
Analogically, logic state coverage represents the \emph{covered area} of a logic
space.

>>>>

  In this work, we propose \emph{logic state coverage} as a proxy metric to
  count observed interesting behaviors. A logic state is a set of satisfied
  branches during one execution, where logic state coverage is the count of
  \emph{individual} observed logic states during a fuzzing campaign. A logic
  state distinguishes less repetitive (i.e., more interesting) behaviors in a
  finer granularity, making the amount of logic state coverage reliably 
  proportional to the number of observed interesting behaviors.
  % 
  We implemented logic state coverage using a bloom filter and reevaluated
  \textsc{FuzzBench} and indecisive-to-subjective fuzzer comparisons in
  hypervisor and grammar fuzzing. 
  % TODO: (The result summary will come here)

@@

As it is challenging to directly count interesting program behaviors when
fuzzing produces hundreds of them every second, we propose \emph{logic state
coverage} as a feasible proxy metric of interesting behaviors. A logic state is
a set of all satisfied branches during one execution, where logic state coverage
is the \emph{individual count} of all observed distinct logic states during a
fuzzing campaign. Logic states represent more \emph{interesting} behaviors
(i.e., relating to unknown abnormal behaviors) in a finer granularity, meaning
that observing more interesting behaviors is reliably reflected to larger
coverage.

@@

\pp{New vs. unknown behaviors}
%
We first make a clear distinction between \emph{new} and \emph{unknown} 
behaviors. \emph{New} behaviors are, as the name suggests, simply the behaviors 
that have not been observed until now. On the other hand, \emph{unknown}
behaviors are a subset of the \emph{new} behaviors that are \emph{not
equivalent} to any observed behaviors. In particular, an abnormal behavior is
\emph{unknown} if it is not equivalent to any observed abnormal behaviors (e.g.,
a unique root cause). Formally, if $\mathrm{U}$ and $\mathrm{N}$ are the sets of
\underline{\textbf{u}}nknown and \underline{\textbf{n}}ew behaviors,
$\mathrm{U} \subset \mathrm{N}$.

\pp{Formal justification}
%
The approach of untargeted fuzzing can be formally described as follows. Upon
executing a mutated input, let $P(\mathrm{N})$ be the probability of observing a
\underline{\textbf{n}}ew behavior as a result of execution. Then, the
probability of discovering \underline{\textbf{u}}nknown
\underline{\textbf{a}}bnormal behavior $P(\mathrm{UA}) := P(\mathrm{U} \cap
\mathrm{A})$ is
%
\begin{align}
  P(\mathrm{UA}) 
      &= P(\mathrm{N} \cap \mathrm{UA}) \tag{by $\mathrm{U} \subset \mathrm{N}$} \\
      &= P(\mathrm{N}) \times P(\mathrm{UA} \mid \mathrm{N}) \tag{by chain rule}. 
\end{align}

@@

\pp{Interestingness in practice}
%
As the interestingness $P(\mathrm{UA} \mid \mathrm{N})$ is the expectation of an
\emph{unobserved} behavior, there is no direct way to measure interestingness. 
%
However, the practice of untargeted fuzzing suggests that
the resulting program behavior is unlikely to be \emph{unknown abnormal} if the
mutation base is highly repetitive, as fuzzers commonly incorporate corpus
minimization \cite{todo} and branch hit count buckets \cite{todo}.
%
The rationales are: (i) if a fewer repetition was normal, more repetition is
also likely normal (i.e., unlikely \emph{abnormal}) and (ii) even if it was
abnormal, it may correspond to pre-observed less-repetition counterparts (i.e.,
unlikely \emph{unknown}).

@@

\pp{Logic state coverage}
%
\emph{Logic state coverage} is \emph{the count of observed distinctive logic
states during fuzzing}, where a logic state is called \emph{covered} when
fuzzing observed a program behavior that belongs to such a logic state (i.e.,
triggering all and only the branches in the logic state). 
%
Analogically, logic state coverage represents the \emph{covered area} of a logic
space.
